---
title: "Infectivity_DCV_FHV_Figure_3"
author: "Angelique Asselin"
date: "2023-09-18"
output: html_document
---


## Experiment Description 
Virus: DCV and FHV 
This document contains the analysis comparing the infectivity of virus grown in the presence or absence of Wolbachia. 
Overal to measure the impact of Wolbachia on the infectivity of DCV and FHV. Infectivity was measured as the number of infectious virus /total virus 


## Approach

Steps
 1. Transformation (log scale) DONE
 2. Check assumptions DONE
 3. t-test DONE
 4. GLM DONE
 5. Compare outputs DONE 
 6. Presentation DONE
 
# Preparation 
 
```{r Preparation loading libraries}
library(tidyverse)
library(ggplot2)
library(ggdensity)
library(ggpubr)
library(moments)
library(car)
library(rstatix)
```

## Importing Data 
sample        = Sample number
IU_mL         = Measure of concentration of infectious particles (IU/ml)
treatment     = Whether the progeny virus was grown in flies with or without Wolbachia W- W+ 
virus         = The type of progeny virus DCV or FHV
biol_rep      = The biological replicate
total_virus   = Concentration of total virus (Total/ml)

```{r Importing data}
# Importing measures of infectious particles and total particles
data <- as.data.frame(read_csv("Infectivity_DCV_FHV_Figure_3_data.csv"))
```


## Transformation 
Transform the data because it is not normally distributed. Log10.
Transform the total and infectious virus before taking the ratio. 
Infectivity = log10(IU)/log10Total
 

```{r 1. Transformation}

# log10 transformation
data <- data %>% 
  mutate(log_total_virus = log10(total_virus)) %>% 
  mutate(log_IU_ml = log10(IU_mL)) %>% 
  mutate(log_infectivity_ratio = log_IU_ml/log_total_virus) %>% 
  mutate(infectivity_ratio = IU_mL/total_virus)

data_1 <- data %>%
  group_by(biol_rep) %>%
  mutate(average_log_infectivity_ratio = mean(log_infectivity_ratio)) %>% 
  mutate(average_infectivity_ratio = mean(infectivity_ratio))

# Collapsing the dataframe so there is only one point for each biol rep that is the average of two tech reps 

c <- seq(from = 1, by =2, length.out = 12)
data_2 <- data_1[c,]

# Formatting
data_2 <-data_2 %>%  
  mutate(average_log_infectivity_ratio = as.double(average_log_infectivity_ratio)) %>% 
  mutate(treatment = as.factor(treatment)) %>% 
  mutate(virus = as.factor(virus))

# Changing the names of the treatments 
# Changing it so that it is W- for Tet and W+ for Wol 
data_2$treatment <- gsub("tet", "W-", data_2$treatment)
data_2$treatment <- gsub("wol", "W+", data_2$treatment)
data_2$treatment <- factor(data_2$treatment, levels = c("W-", "W+"))
data_2$virus <- factor(data_2$virus, levels = c("DCV", "FHV"))

# Subset data based on the virus 
data_DCV <- data_2 %>% 
  subset(virus == "DCV")

data_FHV <- data_2 %>% 
  subset(virus == "FHV")

# Subset the data based on the Wolbachia status 

data_DCV_no_Wol <- data_DCV %>% 
  subset(treatment == "W-")
  
data_DCV_Wol <- data_DCV %>% 
  subset(treatment == "W+")

data_FHV_no_Wol <- data_FHV %>% 
  subset(treatment == "W-")
  
data_FHV_Wol <- data_FHV %>% 
  subset(treatment == "W+")

# Calculating the infectivity 

DCV_no_Wol_infect <- (sum(data_DCV_no_Wol$average_infectivity_ratio))/3
DCV_Wol_infect <- (sum(data_DCV_Wol$average_infectivity_ratio))/3
FHV_no_Wol_infect <- (sum(data_FHV_no_Wol$average_infectivity_ratio))/3
FHV_Wol_infect <- (sum(data_FHV_Wol$average_infectivity_ratio))/3

# Calculating the number of total particles to 1 infectious particle 
DCV_no_Wol_number_infect_total <- 1/DCV_no_Wol_infect
DCV_Wol_number_infect_total <- 1/DCV_Wol_infect
FHV_no_Wol_number_infect_total <- 1/FHV_no_Wol_infect
FHV_Wol_number_infect_total <- 1/FHV_Wol_infect

# Creating a table with these results 
total_to_IU <- rbind(DCV_no_Wol_number_infect_total,DCV_Wol_number_infect_total,FHV_no_Wol_number_infect_total,FHV_Wol_number_infect_total)
total_to_IU <- round(total_to_IU, digits = 0)

Virus <-  c("DCV W-", "DCV W+", "FHV W-", "FHV W+")

infect_number <- as.data.frame(total_to_IU) 
infect_number <- infect_number %>% 
  rename( "Total Virus to 1 infectious particle" = "V1") %>% 
  tibble::rownames_to_column( "Delete") %>% 
  cbind(Virus) %>% 
  select(Virus, "Total Virus to 1 infectious particle", -Delete)

```

## 2. Assumptions
This confirms that logtransforming the data was appropriate. 
Because there are two viruses here I think that they might have a biphasic distribution (therefore check for normality grouping by virus?)

Measuring skewness 
Examples of distributions with  skewness include the following.

skewness = 0: A normal distribution or symmetric distribution 
skewness = less 1 :A half-normal distribution 
skewness = 2: An exponential distribution has a 
skewness = positive: A lognormal distribution 

Results 
Both viruses considered together 
Un-transformed skewness = 1.15 
Transformed skewness = -0.314

DCV
Un-transformed skewness =  -0.4489619
Transformed skewness =-0.2971849

FHV
Un-transformed skewness =  0.6235217
Transformed skewness = -0.6460759

Conclusions 
It appears that the entire dataset has a lognormal distribution that is normalised (reduced to close to three and defs below 1 when log10 transformed)
Continue with the log10 transformation unless otherwise decided. 

```{r 2. Assumptions}
## Plotting distribution of data ## 
# Normality of un-transformed data
ggdensity(data_2, x = "average_infectivity_ratio", fill = "grey")+
  stat_overlay_normal_density(color = "red", linetype = "dashed")

# Normality of transformed data 
ggdensity(data_2, x = "average_log_infectivity_ratio", fill = "grey")+
  stat_overlay_normal_density(color = "red", linetype = "dashed")

# Normality of un-transformed data SORTED BY VIRUS
ggdensity(data_2, x = "average_infectivity_ratio", fill = "grey", facet.by = "virus")+
  stat_overlay_normal_density(color = "red", linetype = "dashed")

# Normality of transformed data SORTED BY VIRUS
ggdensity(data_2, x = "average_log_infectivity_ratio", fill = "grey", facet.by = "virus")+
  stat_overlay_normal_density(color = "red", linetype = "dashed")

## Formally assess normality ## 
skewness(data_2$average_infectivity_ratio, na.rm = TRUE) # 1.15 
skewness(data_2$average_log_infectivity_ratio, na.rm = TRUE) # -0.314

skewness(data_DCV$average_infectivity_ratio, na.rm = TRUE) # -0.44
skewness(data_DCV$average_log_infectivity_ratio, na.rm = TRUE) # -0.29

skewness(data_FHV$average_infectivity_ratio, na.rm = TRUE) # 0.623
skewness(data_FHV$average_log_infectivity_ratio, na.rm = TRUE) # -0.646
```
## 3. T-test

Results 
```{r 3. T-test}
#DCV 
t_test_DCV <- t.test(data = data_DCV, 
                 average_log_infectivity_ratio ~ treatment)

#FHV 
t_test_FHV<- t.test(data = data_FHV, 
                 average_log_infectivity_ratio ~ treatment)
```

## 4. GLM 

 Results 
 
 * 3 df with no interaction 
 * 4 df with interaction 
 * no interaction was a better model (AIC smaller)
 * I don't think there is really a difference in the model (deviance ANOVA = 0 )
 * Use the model 2
 
 
```{r 4. GLM }
## GLM ## 
# With interaction 
model_1 <- glm(formula = average_log_infectivity_ratio ~ treatment * virus,
             family = binomial, 
             data = data_2)
summary(model_1)

# Without interaction
model_2 <- glm(formula = average_log_infectivity_ratio ~ treatment + virus,
             family = binomial, 
             data = data_2)
summary(model_2)

# Best model AIC 
AIC(model_1, model_2)

# Comparing models ANOVA
anova(model_1, model_2)

```

# 4. GLM Assumptions

1. Homogeneity of variances (PASS)
2. Normality (PASS)
3. Outliers (PASS)


```{r 4. GLM Assessing Assumptions}

# 1. Homogeneity of variances
plot(model_2, 1)
leveneTest(average_log_infectivity_ratio ~ 
             virus, 
           data = data_2)

# 2. Normality
plot(model_2, 2)
shapiro_test(residuals(model_2))

# 3. Outliers 
plot(model_2, 1)
outlierTest(model_2)
```


# 5. Presenting the results 

* error bars = SE 
* bar is the mean 

```{r 5. Presenting the results}

summary_data <- data_2 %>%
  group_by(treatment, virus) %>%
  summarize(
    Mean = mean(average_infectivity_ratio),
    SE = sd(average_infectivity_ratio) / sqrt(n())
  )

# Graphing the data 
infectivity <-  ggplot(data_2) +
  geom_jitter(aes(x = interaction(treatment,virus), 
                  y = average_infectivity_ratio, 
                  col = treatment), 
              size = 5, 
              position = position_jitter(width = 0.1, height = 0.0001), 
              show.legend = FALSE)  + 
  scale_color_manual(values = c("black", "red")) +
  scale_shape_manual(values = c(16, 1)) +
  stat_summary(
    data = summary_data,
    aes(x = interaction(treatment, virus),
        y = Mean, 
        group = virus),
    fun.data = mean_se,
    geom = "crossbar",
    width = 0.2,  # Adjust the width of the bars
    col = c("black", "red", "black", "red" )  # Color of the bars
  ) +
  # Adjust the appearance of summary bars and error bars
  geom_errorbar(
    data = summary_data,
    aes(x = interaction(treatment,virus),
        ymin = Mean - SE, 
        ymax = Mean + SE, 
        group = virus),
    width = 0.1,  # Width of error bars
    col = c("black", "black", "red", "red" )  # Color of error bars
  ) + 
  theme_classic() +
  labs(x = "", 
       y = expression(bold("Infectivity ratio"))) +  # Use "\n" for line break
  theme(plot.margin = unit(c(3, 1, 1, 2), "lines"),  # adjust plot margins
        axis.title = element_text(size = 30, face = "bold", colour = "black"),  # center justify axis title text
        axis.title.y = element_text(size = 30, face = "bold", colour = "black", hjust = 0.5, vjust = 0.1),  # center text on the Y-axis
        axis.text = element_text(size = 30, face = "bold", colour = "black"),  # increase size of axis tick labels 
        plot.title = element_text(size = 30 )) +# increase size of plot title text 
  coord_cartesian(ylim = c(0, 0.005)) + # sets the maximum y-axis 
  scale_x_discrete(labels = c("W-\n DCV", "W+\n DCV", "W-\n FHV", "W+\n FHV")) +
  scale_y_continuous(breaks = c(0,0.0025, 0.005))

```


```{r Export plots run outside of RMD}

# Exporting the specific infectivity summary table
write.csv(infect_number, file = "Infectivity_DCV_FHV_Figure_3_Output_Specific_Infectivity.txt")

# Exports T-test output 
# DCV 
sink("Infectivity_DCV_FHV_Figure_3_Output_DCV_T_test.txt") 
print(t_test_DCV) 
sink() # closes the connection **CRITICAL**

# FHV
sink("Infectivity_DCV_FHV_Figure_3_Output_FHV_T_test.txt") 
print(t_test_FHV) 
sink() # closes the connection **CRITICAL**

# Exports GLM
# DCV and FHV model 
sink("Infectivity_DCV_FHV_Figure_3_Output_GLM.txt") # change file name
print(model_2) # prints coxme output to txt file
sink() # closes the connection **CRITICAL**

# DCV and FHV model SUMMARY 
sink("Infectivity_DCV_FHV_Figure_3_Output_GLM_summary.txt") # change file name
print(summary(model_2)) # prints coxme output to txt file
sink() # closes the connection **CRITICAL**

# Exporting figure
today <- Sys.Date()
date <- format(today, format = "%y%m%d")
name <- paste(date, "Figure", "3",  sep = "_")
pdf_name <- paste(name, "pdf", sep = ".")
ggsave(filename = pdf_name,
       plot = infectivity, 
       width = 8, 
       height = 4)

```

