---
title: "Infectivity_DCV_FHV_04"
author: "Angelique Asselin"
date: "2023-09-18"
output: html_document
---

## Editing log 
18/09/23 - Making a more comprehensive log of the analysis that was conducted. 

## QUESTIONS

* Do I need to facet by the viral type or because it is a combined assay where the values are normalised should it just be one test for normality and skewness? (if separated into the two virus treatments then it only looks at like 3 data points so can you really determine a distribution using that amount of information?)

## Experiment Description 
Virus: DCV and FHV 
Assay: 04 measuring infectivity ratio 

## Meeting with Andrew to discuss analysis 

Infectivity 04 quantifying the impact of Wolbachia on the infectivity of progeny particles 

	- Statistical analysis: Put results from infectivity of two viruses into the same model (would this be mapping the difference of the ratio between Wolbachia and Tet flies rather than the actual infectivity ratios?) 
	- What would you recommend?
- IR ~ Wolbachia + Virus  
- IR ~ Wolbachia * Virus 
	- Compare those two models see which one is better

IR (Tet + DCV) = 0.01 
IR (Wol +DCV) = 0.01 + x1 
IR (Tet +FHV) = 0.01 + x2 
IR (Wol+ FHV) = 0.01 + x1 + x2 + x3 

 Analysis 
	- GLM for ratio (0,1) 
	- My ratio will always be between 0 and 1 because you can't have more IU than virus particles 
		- 1. Beta regression (if not dealing with whole numbers) 
		- 2. Logistic regression GLm (y~ x, family = binomial) 
			- Whole numbers 
			- Andrew suggested this one if i have whole numbers 
			- Google has the answers 
			- let him know if I need a hand 
			- I think I want multinomial 

	1. Transform the data because it is not normally distributed (lognormal?) (transform the data before taking the ratio (because if you take the log of the ratio it does weird things and it is no longer 0-1) 
	2. Conduct a T-teston  the viruses separately ( = 5df) 
	3. GLM (with and without an interaction) (= 8df) do an ANOVA (also check AIC) to figure out if these are the same things. 
	4. See if the output is the same  
	5. Choose which to present in the paper the simple version and then have the complicated version in the appendix

## Description of RMD

This document is a log of the analysis conducted for the infectivity assay. 
This looks at the infectivity ratio infectious/total virus particles of 
DCV and FHV grown in the presence and absence of Wolbachia. 

## Statistical Approach 

 GLM for a ratio (0,1)
Logitic regression GLM( y ~x, family = binomial)
IR ~ Wolbachia + Virus (No interaction)
IR ~ Wolbachia * Virus 

Steps
 1. Transformation (log scale) DONE
 2. Check assumptions DONE
 3. t-test DONE
 4. GLM DONE
 5. Compare outputs DONE 
 6. Choose presentation 
 
 
 
```{r Install Packages}
install.packages("ggdensity")

```

```{r Preparation}
library(tidyverse)
library(ggplot2)
library(ggdensity)
library(ggpubr)
library(moments)

```
## Transformation 
Transform the data because it is not normally distributed. Log10.
Transform the total and infectious virus before taking the ratio. 
log10(IU)/log10Total
If you take the log of the ratio it does weird things and it is no longer 0-1.  

```{r 1. Transformation}
# Reading the data (from laptop)
data <- read.csv("/Users/angeliqueasselin/Documents/GitHub/infectivity_04/analysed/230131_Analysed_FHV_DCV.csv")


# Transforming the data 
c <- rep(1e+12, times =24) 

# log10 transformation
data <- data %>% 
  mutate(total_virus = c) %>% 
  mutate(log_total_virus = log10(total_virus)) %>% 
  mutate(log_IU_ml = log10(IU_mL)) %>% 
  mutate(log_infectivity_ratio = log_IU_ml/log_total_virus)

data_1 <- data %>%
  group_by(biol_rep) %>%
  mutate(average_log_infectivity_ratio = mean(log_infectivity_ratio)) 

# Collapsing the dataframe so there is only one point for each biol rep that is the average of two tech reps 

c <- seq(from = 1, by =2, length.out = 12)
data_2 <- data_1[c,]

# Formating

data_2 <-data_2 %>%  
  mutate(average_log_infectivity_ratio = as.integer(average_log_infectivity_ratio))

# Changing the names of the treatments 
# Changing it so that it is W- for Tet and W+ for Wol 
data_2$treatment <- gsub("tet", "W-", data_2$treatment)
data_2$treatment <- gsub("wol", "W+", data_2$treatment)

# Subset data based on the virus 
data_DCV <- data_2 %>% 
  subset(virus == "DCV")

data_FHV <- data_2 %>% 
  subset(virus == "FHV")

```

## 2. Assumption test 
This confirms that logtransforming the data was appropriate. 
Because there are two viruses here I think that they might have a biphasic distribution (therefore check for normality grouping by virus?)

Measuring skewness 
Examples of distributions with  skewness include the following.

skewness = 0: A normal distribution or symmetric distribution 
skewness = less 1 :A half-normal distribution 
skewness = 2: An exponential distribution has a 
skewness = positive: A lognormal distribution 

Results 
Both viruses considered together 
Un-transformed skewness = 1.15 
Transformed skewness = -0.314

DCV
Un-transformed skewness =  -0.4489619
Transformed skewness =-0.2971849

FHV
Un-transformed skewness =  0.6235217
Transformed skewness = -0.6460759

Conclusions 
It appears that the entire dataset has a lognormal distribution that is normalised (reduced to close to three and defs below 1 when log10 transformed)
Continue with the log10 transformation unless otherwise decided. 

```{r 2. Assumptions}
## Plotting distribution of data ## 
# Normality of un-transformed data
ggdensity(data_2, x = "average_infectivity_ratio", fill = "grey")+
  stat_overlay_normal_density(color = "red", linetype = "dashed")

# Normality of transformed data 
ggdensity(data_2, x = "average_log_infectivity_ratio", fill = "grey")+
  stat_overlay_normal_density(color = "red", linetype = "dashed")

# Normality of un-transformed data SORTED BY VIRUS
ggdensity(data_2, x = "average_infectivity_ratio", fill = "grey", facet.by = "virus")+
  stat_overlay_normal_density(color = "red", linetype = "dashed")

# Normality of transformed data SORTED BY VIRUS
ggdensity(data_2, x = "average_log_infectivity_ratio", fill = "grey", facet.by = "virus")+
  stat_overlay_normal_density(color = "red", linetype = "dashed")

## Formally assess normality ## 
skewness(data_2$average_infectivity_ratio, na.rm = TRUE) # 1.15 
skewness(data_2$average_log_infectivity_ratio, na.rm = TRUE) # -0.314

skewness(data_DCV$average_infectivity_ratio, na.rm = TRUE) # -0.44
skewness(data_DCV$average_log_infectivity_ratio, na.rm = TRUE) # -0.29

skewness(data_FHV$average_infectivity_ratio, na.rm = TRUE) # 0.623
skewness(data_FHV$average_log_infectivity_ratio, na.rm = TRUE) # -0.646
```
## 3. T-test

Results 
```{r 3. T-test}
#DCV 
t_test_DCV <- t.test(data = data_DCV, 
                 average_log_infectivity_ratio ~ treatment)

#FHV 
t_test_FHV<- t.test(data = data_FHV, 
                 average_log_infectivity_ratio ~ treatment)
```

```{r 2. T-test Exporting results (RUN OUTSIDE RMD)}
 # Exports T-test output 

# DCV 
sink("results/230918_IFB_04_t_test_output.txt") # change file name
print(t_test_DCV) # prints coxme output to txt file
sink() # closes the connection **CRITICAL**

# FHV
sink("results/230918_IFC_04_t_test_output.txt") # change file name
print(t_test_FHV) # prints coxme output to txt file
sink() # closes the connection **CRITICAL**

```
 
 
 ## 4. GLM 

 Results 
 
 * 3 df with no interaction 
 * 4 df with interaction 
 * no interaction was a better model (AIC smaller )
 * I don't think there is really a difference in the model (deviance ANOVA = 0 )
 
 Future 
 * Use the model 2
```{r 4. GLM }
## GLM ## 
# With interaction 
model_1 <- glm(formula = average_log_infectivity_ratio ~ treatment * virus,
             family = binomial, 
             data = data_2)
summary(model_1)

# Without interaction
model_2 <- glm(formula = average_log_infectivity_ratio ~ treatment + virus,
             family = binomial, 
             data = data_2)
summary(model_2)

# Best model AIC 
AIC(model_1, model_2)

# Comparing models ANOVA
anova(model_1, model_2)

```
```{r 3. GLM exporting results(RUN OUTSIDE RMD) }
# Exports GLM

# DCV and FHV model 
sink("results/230918_IFB_IFC_04_GLM_output.txt") # change file name
print(model_2) # prints coxme output to txt file
sink() # closes the connection **CRITICAL**

# DCV and FHV model SUMMARY 
sink("results/230918_IFB_IFC_04_GLM_output.txt") # change file name
print(summary(model_2)) # prints coxme output to txt file
sink() # closes the connection **CRITICAL**
```

## 5. Presenting the results 

* error bars = SE 
* bar is the mean 

```{r 5. Presenting the results}

summary_data <- data_2 %>%
  group_by(treatment, virus) %>%
  summarize(
    Mean = mean(average_infectivity_ratio),
    SE = sd(average_infectivity_ratio) / sqrt(n())
  )

# Graphing the data 
infectivity <- ggplot(data_2) +
  geom_jitter(aes(treatment, 
                  average_infectivity_ratio, 
                  col = treatment, 
                  shape = virus), 
              size = 5, 
              position = position_jitter(width = 0.1), 
              show.legend = FALSE)  + 
  scale_color_manual(values = c("black", "red")) +
  scale_shape_manual(values = c(16, 1)) +
  stat_summary(
    data = summary_data,
    aes(x = treatment, y = Mean, group = virus),
    fun.data = mean_se,
    geom = "crossbar",
    width = 0.2,  # Adjust the width of the bars
    col = c("black", "red", "black", "red" )  # Color of the bars
  ) +
  # Adjust the appearance of summary bars and error bars
  geom_errorbar(
    data = summary_data,
    aes(x = treatment, ymin = Mean - SE, ymax = Mean + SE, group = virus),
    width = 0.1,  # Width of error bars
    col = c("red", "red", "black", "black" )  # Color of error bars
  ) + 
  theme_classic() +
  labs(x = "", 
       y = expression(bold("Log Normalised infectivity ratio"))) +  # Use "\n" for line break
  theme(plot.margin = unit(c(3, 1, 1, 2), "lines"),  # adjust plot margins
        axis.title = element_text(size = 30, face = "bold", colour = "black"),  # center justify axis title text
        axis.title.y = element_text(size = 30, face = "bold", colour = "black", hjust = 0.5, vjust = 0.1),  # center text on the Y-axis
        axis.text = element_text(size = 30, face = "bold", colour = "black"),  # increase size of axis tick labels 
        plot.title = element_text(size = 30 ))  # increase size of plot title text 
```


```{r Export plots run outside of RMD}
# Exporting the results 

today <- Sys.Date()
date <- format(today, format = "%y%m%d")

# Saving plot
name <- paste(date, "IFB","IFC", "infectivity", "04", sep = "_")
pdf_name <- paste(name, "pdf", sep = ".")
ggsave(filename = pdf_name,
       path = "figures",
       plot = infectivity, 
       width = 15, 
       height = 7.5)

```

